```{r}
#source("requirements.R")
source("definitions.R")
```
# Maximum cross correlation frequency per country. 
News evidence suggests this should be around 21 or at least 15. It was much less on the data of Jan- April 2020

```{r}
findMaxCCF(myPSCR = "Italy")
findMaxCCF(myPSCR = "Hubei,China")
findMaxCCFs(myPSCR = "Taiwan*")
```
Italy 28, Hubei 19,  and Taiwan 26 days, in the data until 2020-09-22. 

## Recovered lagging Confirmed by...
```{r}
rclags <- findMaxCCFs("new_recovered","new_confirmed")
rclags <- rclags[!is.nan(rclags$cor), ]  #see comment below about NAN and NA
hist(rclags$lag, plot = TRUE, breaks = 20)
```
```{r}
if (every( rclags[rclags$lag >= 0,"lag"], is.numeric)) 
   rclags[rclags$lag >= 0,][["lag"]] %>% median
```
It seems that as of 2020-09-22, the median is now 14 days, with the bulk of the lags are between 10 to 20 days, with secondary peaks at zero, and at 26-27 days. Some 8 countries still exhibit negative lags. 

In April 2020, it seemed 21 days was the mode time to recover since being confirmed infected for Taiwan*. By end of September, it became 26 days for Taiwan. 

Singapore tested the first 100 patients, and all recover by the 31st day. Some the first day! 
https://towardsdatascience.com/visual-notes-from-singapores-first-100-fully-recovered-covid-19-patients-aad7f2e1d0a0
majority of patients among the first 100 fully recovered cases were confirmed within 14 days of the reported onset of symptom

Sweden claims patients all recover by the 6th week, 42 days. Ref(?) 

```{r}
rclags[order(rclags$lag,decreasing = TRUE),][1:20,]
```

```{r}
rclags[order(rclags$lag,decreasing = FALSE),][1:15,]
```

## Recovered lagging Deaths by ...
```{r}
verbose = 1
rdlags <- findMaxCCFs("new_recovered","new_deaths")
rdlags <- rdlags[!is.nan(rdlags$cor),]
hist(rdlags$lag, plot = TRUE,breaks = 20)
```
It seems recovery lags deaths , but a lot of regions have lag around zero still. 

```{r}
#rdlags[rdlags$lag<=5&rdlags$lag>=0,"PSCR"]
rdlags[rdlags$lag >= 0,][["lag"]] %>% median
```
```{r}
rdlags[rdlags$lag > 5,][["PSCR"]]
```
```{r}
rdlags[rdlags$lag > 10,][["PSCR"]]
rdlags[order(rdlags$lag,decreasing = TRUE),][1:20,]
```
## Deaths lagging Confirmed by ...
```{r}
dclags <- findMaxCCFs("new_deaths","new_confirmed")
dclags <- dclags[!is.nan(dclags$cor),]
hist(dclags$lag, plot = TRUE,breaks = 20)
dclags[dclags$lag >= 0,][["lag"]] %>% median
```
```{r}
dclags[order(dclags$lag,decreasing = TRUE),][1:20,]
```
## Discussion: 
So I suggest taking lags 42 and 36 (i.e. assume people die within 6 days if they die, which (in April) was the median of positive rd lags. 
By September, the median DC lag equals 6. 

Note we do not take the median positive RC lag, which equals 14 (JHH data until 2020-05-18), 
if we take a smaller lag between r and d, then d might start rising faster than c, which leads to r decreasing over time and that is not possible.the New York,US data we have imputed suffers from this idiosyncracy.  )

bug in CCF output: instead of cor=NA, and lag=NA, we obtain cor =NAN, lag =-28 when var1 is missing. The Documentation says it should be NA. So we do filter NAN insteead of NA. If future versions give an error, check the NA/NAN issue. 

```{r}
LAGDC; LAGRD;LAGRC
```
