#####Making GIFS
Animate .gif images in R / ImageMagick
November 21, 2010
By markheckmann


Yesterday I surfed the web looking for 3D wireframe examples to explain linear models in class. I stumbled across this site where animated 3D wireframe plots are outputted by SAS.  Below I did something similar in R. This post shows the few steps of how to create an animated .gif file using R and ImageMagick. Here I assume that you have ImageMagick installed on your computer. As far as I know it is also possible to produce animated .gif files using R only, e.g. with write.gif() from the caTools package. But using ImageMagick is straighforward, gives you control over the conversion and .gif production and is the free standard program for conversion.

First a simple countdown example. To be sure not to overwrite anything I will create a new folder and set the working directory to the new folder.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 example 1: simple animated countdown from 10 to "GO!".
```{r}
dir.create("examples")
setwd("examples")
png(filename = "example%02d.png", width = 200, height = 200)
for (i in c(10:1, "G0!")) {
plot.new()
text(.5, .5, i, cex = 6)
}
dev.off()
```
 convert the .png files to one .gif file using ImageMagick. 
 The system() function executes the command as if it was done
 in the terminal. The -delay flag sets the time between showing
 the frames, i.e. the speed of the animation.
```{r}
system("convert -delay 80 *.png example_1.gif")
``` 
this works only if you have image magick. 
```{r echo=TRUE}
myGraph <- 'graph3Dard_fia'
myBy = '1 day'
makeHistoryGraph(JHH, regios$Vincent, graph = myGraph, 
            toDates = seq.Date(as.Date('2020-01-22'), Sys.Date(), by = myBy),
            myFolderType = myBy)
system("convert -delay 80 " % % myPlotPath %//% graph %//% '1' %//%"*.png "
       % % graph %#%".gif")
```
you can also use a package: 
```{r} 
devtools::install_github("dgrtwo/gganimate")# and install image_magick. 
# see http://www.ggtern.com/2017/07/23/version-2-2-1-released/
# add  frame = Date  to aes of ggplot, then
# gganimate(myplot, "filename.gif" )  # or mp4, html, swf. 
# p3 <- ggplot(gapminder, aes(gdpPercap, lifeExp, frame = year)) +
#geom_path(aes(cumulative = TRUE, group = country)) +
#  scale_x_log10() +
#  facet_wrap(~continent)
```
