---
title: "Graphs"
author: "VF"
date: "9 July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Resulting graphs and cleaned and prepared data can be found at 
https://drive.google.com/drive/folders/1Yo0IW4awCIvWMP6jYVKpn3kLgEolse0e?usp=sharing
#for publication 
```{r cleanup load latest graph definitions and data, echo=FALSE}
rm(list = setdiff(ls(), c('ECDC0', 'JHH0', 'JHH', 'ECDC', 'JHHRegios', 'ECDCRegios','testing')))
source('definitions.R')
options(warn = 0)
if (!exists('JHH')) source('loadData.R') else 
  if (max(JHH$Date) < Sys.Date() - 1) source('loadData.R')  
```
Graph active_imputed, recovered, deaths, and confirmed for some selected territories, 
based on JHH data:
```{r demo graph}
graphit(JHH, regios$Vincent, xvar = 'Date', 
        yvars = c('active_imputed','recovered_imputed', 'deaths','confirmed'),
        facet = "PSCR", logy = TRUE)
```
Show table outputs: 
```{r}
#show some data table output
#latest numbers
JHH[JHH$Date == max(JHH$Date),'Date'][1,1]
JHH %>% filter(Date == max(Date)) %>% filter(!is.nan(new_active_rate)) %>%
  select(PSCR,confirmed, active_imputed, deaths, new_confirmed, new_active_rate, active_imputed_growthRate, 
         active_imputed_p_M) %>% arrange(new_active_rate) %>% tail(10)
JHH[JHH$Date == max(JHH$Date) & JHH$PSCR %in% 
      c('EU','World','New York,US',"Kazakhstan",'Belgium','Spain','US','Netherlands','Europe',
        'Germany','France','Africa','Iran','Russia','Brazil'),
    c('PSCR','confirmed','active_imputed','deaths','new_confirmed',
      'new_active_rate', 'active_imputed_growthRate','active_imputed_p_M') ]  %>%
  arrange(new_active_rate)

```
Who overtakes us based the JHH data set on latest day data? 
```{r} 
map_dfc(c('Kazakhstan','Belgium','Netherlands','Sweden'),function(x) overtakeDays_df(JHH,x,who = 'theyme',lastDays = 1))
```
And on 7 day averages in JHH? 
```{r}
map_dfc(c('Kazakhstan','Belgium','Netherlands','Sweden'),function(x) overtakeDays_df(JHH,x,who = 'theyme',lastDays = 7))
```

And according to ECDC? 
```{r}
map_dfc(c('Kazakhstan','Belgium','Netherlands','Sweden'),function(x) overtakeDays_df(ECDC,x,who = 'theyme',lastDays = 7))
```
Who is more affected? 
```{r}
graph3Dard_fia(ECDC, c('Kazakhstan','Belgium','Netherlands','Sweden'))
```
Who do we overtake?
```{r}
map_dfc(c('Kazakhstan','Belgium','Netherlands','Sweden'),function(x) overtakeDays_df(JHH,x,who = 'Ithem',lastDays = 7))
map_dfc(c('Kazakhstan','Belgium','Netherlands','Sweden'),function(x) overtakeDays_df(ECDC,x,who = 'Ithem',lastDays = 7))
```
Some other countries now: 
```{r}
graph3Dard_fia(JHH, c('Germany','France','United Kingdom'))
map_dfc(c('Germany','France','United Kingdom'), function(x) overtakeDays_df(JHH,x,who = "theyme", lastDays = 7))
map_dfc(c('Germany','France','United Kingdom'), function(x) overtakeDays_df(JHH,x,who = "Ithem", lastDays = 7))

```
For the New York state, Indonesia, Peru, India: 
```{r}
graph3Dard_fia(JHH, c('New York,US','Indonesia','Peru','India'))
map_dfc(c('New York,US','Indonesia','Peru','India'),function(x) overtakeDays_df(JHH,x,who = 'Ithem'))
map_dfc(c('New York,US','Indonesia','Peru','India'),function(x) overtakeDays_df(JHH,x,who = 'theyme'))
```
See how the most affected regions in the world are developing: (pop the graph out to a new window and enlarge to see more detail)
```{r}
graph6Dardcra_finyl(JHH, JHHRegios$`JHH World1`)
```
If you want to write the data to disk, make sure to define the correct datapath so that R can write to it. By default it is a subfolder of the scripts folder. Note that the data will also be written to a subfolder of the plots folder. 
``` {r}
dataPath <- './data'
if (!dir.exists(dataPath)) dir.create(dataPath, recursive = TRUE)
writeWithCounters(ECDC,name = "Covid19ECDC")
writeWithCounters(JHH,name = "Covid19JHH") 
```
check what graphs are defined 
```{r}
graphCodes()
myGraphListbyDate
myGraphListbyDay
myGraphNrs
myGraphList
```

